
Exploratory Data Analysis of White Wine Quality
========================================================
#####Jocelyn(Yuan) Li

####**Overview**: 
This tidy data set contains 4,898 white wines with 11 variables on quantifying the chemical properties of each wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

####**Question**:
The scope of this analysis is to analyze Which chemical properties influence the quality of white wines?

####**Variable description**:
1. **Fixed acidity**: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
2. **Volatile acidity**: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste
3. **Citric acid**: found in small quantities, citric acid can add 'freshness' and flavor to wines
4. **Residual sugar**: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and
wines with greater than 45 grams/liter are considered sweet
5. **Chlorides**: the amount of salt in the wine
6. **Free sulfur dioxide**: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine
7. **Total sulfur dioxide**: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
8. **Density**: the density of water is close to that of water depending on the percent alcohol and sugar content
9. **pH**: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
10. **Sulphates**: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant
11. **Alcohol**: the percent alcohol content of the wine
12. **Quality**: output variable (based on sensory data, score between 0 and 10)  


- - -

####**Summary of the Data Set**
```{r}
wine<-read.csv('wineQualityWhites.csv')
dim(wine)
str(wine)
names(wine)
```

- - -

####**Distribution of White Wine Quality**

```{r}

library(ggplot2)
ggplot(aes(x=quality),data=wine)+
  geom_histogram(color='black',fill='#099DD9',binwidth = 0.5)+
  geom_vline(aes(xintercept = median(quality)), color="red", linetype = 1) +
  geom_vline(aes(xintercept = mean(quality)), color="blue", linetype = 1)+
  scale_x_continuous(breaks = seq(3,9,1))+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10))+ggtitle('White Wine Quality')
```


#####For most of the wine in our dataset, quality falls between 5 and 7, the white wine spread shows a normal distribution and have a peak of 6.
 
- - -

####**Correclations Analysis**

```{r}
cor(x=wine[,1:12], y=wine$quality)

library(GGally)
library(corrplot)

theme_set(theme_minimal(20))
set.seed(3000)
ggpairs(wine[sample.int(nrow(wine),1000),])

wine_samp <- wine[,c(1:12)]
corrplot(cor(wine_samp), method = "circle")
```

##### Findinds from the scatterplot matrices
1. PH & fixed.acidity -->Negative correlation
2. Density & quality -->Negative correlation
3. Density & Alcohol -->Negative correlation
4. Density & residual.sugar -->Positive correlation,etc.
5. residual.sugar & alcohol -->Negative correlation

- - -

####**Bivariate Analysis**
#### Alcohol and Quality
```{r echo=FALSE}
#####2. Instead analyzing the quality on all levels, I will create a new variables and cut the quality into 3 range.
library(ggplot2)

wine$quality.bucket <- cut(wine$quality,c(2,5,6,9))
summary(wine$quality.bucket)

ggplot(aes(x=quality.bucket,y=alcohol),data=wine)+geom_boxplot(aes(fill=quality.bucket))+scale_fill_brewer(type = 'qual')+geom_smooth(method='lm', aes(group = 0),color='red')+ggtitle('Alcohol and Quality Relationship')+
  theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10))
```

* * *

#### Density and Quality

```{r echo=FALSE}
library(ggplot2)
ggplot(aes(x=quality.bucket,y=density),data=wine)+geom_boxplot(aes(fill=quality.bucket))+scale_fill_brewer(type = 'qual')+ggtitle('Density and  Quality Relationship')+geom_smooth(method='lm', aes(group = 0),color='red')+ylim(0.99,1.005)+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10),legend.title=element_text(size=12),legend.text=element_text(size=12))

cor.test(wine$density,wine$quality,method = 'pearson')
```

* * *

#### Density and Alcohol
```{r echo=FALSE}
ggplot(aes(x=density,y=alcohol),data=wine)+geom_point()+geom_smooth(method = "lm")+ggtitle('Density and Alcohol Relationship')+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10))+xlim(0.9871,1.01)

cor.test(wine$density,wine$alcohol,method = 'pearson')
```

* * *

#### Density and residual.sugar
```{r echo=FALSE}
ggplot(aes(x=density,y=residual.sugar),data=wine)+geom_point()+geom_smooth(method = "lm")+ggtitle('Density and residual.sugar Relationship')+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10))+xlim(0.9871,1.005)+ylim(0.600,30)

cor.test(wine$density,wine$residual.sugar,method = 'pearson')
```

* * *

#### Density and total.sulfur.dioxide
```{r echo=FALSE}
ggplot(aes(x=total.sulfur.dioxide,y=total.sulfur.dioxide),data=wine)+geom_point()+geom_smooth(method = "lm")+ggtitle('Density and total.sulfur.dioxide Relationship')+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10))

cor.test(wine$density,wine$total.sulfur.dioxide,method = 'pearson')
```

* * *

#### pH and fixed.acidity
```{r echo=FALSE}
ggplot(aes(x=pH,y=fixed.acidity),data=wine)+geom_point()+geom_smooth(method = "lm")+ggtitle('pH and fixed.acidity Relationship')+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10))

cor.test(wine$density,wine$total.sulfur.dioxide,method = 'pearson')
```

##### The strongest relation from bivariate analysis is the negative correlation between density and alcohol.

- - -

#### **Multivariate Plots Section**
#### Density & Alcohol & Wine Quality

```{r echo=FALSE}
ggplot(aes(x=density,y=alcohol),data=wine)+geom_point(aes(color=quality.bucket))+ggtitle('Density & Alcohol & Wine Quality')+xlim(0.98,1.01)+geom_smooth(method = "lm")+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10),legend.title=element_text(size=12),legend.text=element_text(size=12))

```

1. From the graph we can see, higher quality wine tends to have high alcohol and lower density.
2. With same density, higher quality wine tends to have higher alcohol.
3. Also we can find the negative correlation between alcohol and density.

* * *

#### Density & Sugar & Wine Quality
```{r echo=FALSE}
ggplot(aes(x=density,y=residual.sugar),data=wine)+geom_point(aes(color=quality.bucket))+ggtitle('Density & Residual.sugar & Wine Quality ')+xlim(0.985,1.01)+ylim(0,30)+geom_smooth(method = "lm")+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10),legend.title=element_text(size=12),legend.text=element_text(size=12))
```

1. From the graph we can see, higher quality wine tends to have lower residual.sugar and lower density.
2. With same density, higher quality wine tends to have higher residual.sugar
3. Also we can find the positive correlation between residual.sugar and density.

* * *

#### Density & total.sulfur.dioxide & Wine Quality
```{r echo=FALSE}
ggplot(aes(x=density,y=total.sulfur.dioxide),data=wine)+geom_point(aes(color=quality.bucket))+ggtitle('Density & total.sulfur.dioxide & Wine Quality ')+xlim(0.985,1.005)+ylim(0,300)+geom_smooth(method = "lm")+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10),legend.title=element_text(size=12),legend.text=element_text(size=12))
```

1. From the graph we can see, higher quality wine tends to have lower SO2 and lower density.
2. With same density, higher quality wine tends to have lower SO2.
3. Also we can find the positive correlation between 

* * *

#### Density & residual.sugar & alcohol
```{r echo=FALSE}
ggplot(aes(x=density,y=residual.sugar),data=wine)+geom_point(aes(color=alcohol))+ggtitle('Density & Residual.sugar & Alcohol')+xlim(0.985,1.005)+ylim(0,30)+geom_smooth(method = "lm")+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10),legend.title=element_text(size=12),legend.text=element_text(size=12))

```
#####From the graph we can see, With same density, higher volume of alcohol wine tends to have lower sugar.

- - -

#### linear regression 
```{r echo=FALSE}
#####linear regression 
quality_linear=lm(formula = quality~.,data=wine_samp)
quality_linear
summary(quality_linear)
```

#####Due to the correlations between several independent variables (multicollinearity), the linear models doesn't provide accurate results, the adjusted R-squared is small also.


- - -

#### **Final Plots and Summary**

#### Plot One - "Alcohol by quality"

```{r echo=FALSE}
library(ggplot2)
ggplot(aes(x=alcohol),data =wine) + 
  geom_density(aes(fill = quality.bucket),, alpha = 1/3)+
  xlab("Alcohol") + 
  ylab("") +ggtitle("Alcohol by quality")+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10),legend.title=element_text(size=12),legend.text=element_text(size=12))

```

#####Description One
From the density plot we can see the relationship between wine quality and alcohol. Higher quality wine has more alcohol.

* * *

####Plot Two - "Density by quality"
```{r echo=FALSE}
library(ggplot2)
ggplot(aes(x=density),data =wine) + 
  geom_density(aes(fill = quality.bucket),, alpha = 1/3)+
  xlab("Density") + 
  ylab("") +ggtitle("Density by quality")+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10),legend.title=element_text(size=12),legend.text=element_text(size=12))
```

#####Description Two
From the density plot we can see the relationship between wine quality and density, Higher quality wine tends to have less density.

* * *

####Plot Three - "Density by quality"
```{r echo=FALSE}
ggplot(aes(x=density,y=residual.sugar),data=wine)+geom_point(aes(color=alcohol))+ggtitle('Density & Residual.sugar & Alcohol')+xlim(0.985,1.005)+ylim(0,30)+geom_smooth(method = "lm")+theme(plot.title = element_text(size=12,face="bold"),axis.title=element_text(size=12,face="bold"),axis.text =element_text(size=10),legend.title=element_text(size=12),legend.text=element_text(size=12))
```

#### Description Three

##### We've also found some interesting relationship between residual sugar, density and alcohol: sweater wine has more density and wine with the same sweetness has larger volume of alcohol with lower density. 


- - -

#####Reflection
1. Good wine tend to have more alcoho, probably because alcohol kills more residual.sugar or SO2.
3. Good wine tend to have lower density, probably because lower density has lower residual.sugar or SO2.
2. I hope I can take more courses and build a more accurate model to predict the wine quality.

